\documentclass[11pt,letterpaper,oneside]{article}
\usepackage{url}
\begin{document}

\section*{The Data}
For the past year, I have been downloading Craigslist apartment,
rental, sublet and roommate listings, nightly. When a persons posts a
listing, they have the option of putting in an address or
intersection. This allows me to associate latitude and longitude with
a posting. They also have the option of entering, in free text, the
``Specific location'' of the posting.  With these data, I can estimate
how likely that a listing will say that a location is in a
neighborhood (Figure 1).

I have been downloading these data for Chicago since April 2011, and
since December 2011 the following cities: New York, Los Angeles,
Houston, Philadelphia, Phoenix, San Antonio, San Diego, Dallas,
Jacksonville, Indianapolis, San Francisco, Austin, Columbus,
Charlotte, Detroit, El Paso, Memphis, Baltimore, Boston, Seattle,
Washington D.C., Nashville, Denver, Louisville, Milwaukee, Portland,
Las Vegas, Oklahoma City, Albuquerque, Tucson, Fresno, Sacramento,
Kansas City, Atlanta, Colorado Springs, Omaha, Raleigh, Miami,
Cleveland, Tulsa, Minneapolis, Wichita, Knoxville, Asheville. These
data are available through RSS feeds like
\url{http://chicago.craigslist.org/sub/index.rss} or
\url{http://memphis.craigslist.org/roo/index.rss}.

\section*{Privacy Protection Measures}
Going forward, I will take the following steps to protect the privacy
of Craigslist posters. Every month, after identifying duplicates I will 
\begin{enumerate}
\item{randomly perturb locations that were identified by geocoding addresses,}
\item{round the given price to the nearest hundred,}
\item{record the posting data only to the month and year,}
\item{and dispose of the title and text of the posting, and store only
    a list of word pairs and how often they appear in a posting. Word
    pairs that only appear in one posting will not be recorded}
\end{enumerate}

The collection process is completely automated, and this
de-identification will be as well.
\end{document}
